---
title: "ICES VMS datacall quality check report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

<!-- QCTEMPLATE: header -->
```{r setup, eval=TRUE, echo=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   = NA,
               warning   = FALSE,
               message   = FALSE,
               error     = FALSE,
               echo      = FALSE,
               eval      = TRUE,
               dev = "my_png",
               fig.ext = "png",
               fig.align = 'left',
               tab.align = 'left',
               fig.width = 5.5)

options(knitr.kable.NA = '-')

my_png <- function(file, width, height) {
  png(file, width = width, height = height, 
      res = 800, units = "in", pointsize = 20)
}

library(plyr)
library(ggplot2)
library(vmstools)
library(RColorBrewer)
library(doBy)
library(reshape2)
```

```{r load-data}
load("POLYGONS/eurPols.Rdata")
polLand <- fortify(eurPols)
```
<!-- QCTEMPLATE: data -->
<!------------------------------------------------------------------------------
Data handling
---------------------------------------------------------------------------- -->
```{r, echo=FALSE, results='asis'}
#Read in latest submission -->
# NOTE:
#   This chunk of code is replaced when a template is created.

ICES_LE <- read.table('data/ICES_LE_SWE.csv', sep = ',', header = TRUE,
          stringsAsFactors = FALSE, na.strings = 'NULL',
          colClasses = c('character', 'character', 'numeric', 'numeric',
                         'character', 'character', 'character', 'numeric',
                         'character', 'character',
                         'numeric', 'numeric', 'numeric'))
ICES_VE <- read.table('data/ICES_VE_SWE.csv', sep = ',', header = TRUE,
          stringsAsFactors = FALSE, na.strings = 'NULL',
          colClasses = c('character', 'character', 'numeric', 'numeric',
                         'character', 'character', 'character', 'character',
                         'numeric', 'numeric', 'numeric', 'numeric',
                         'numeric', 'numeric', 'numeric', 'numeric'))
```
<!-- QCTEMPLATE: body -->
<!------------------------------------------------------------------------------
Pre-calculations
---------------------------------------------------------------------------- -->


<!--------------------------
Checks for VMS data
------------------------ -->
```{r vms-checks}
#Most recent submission -->
country <- ICES_VE$country[1]
ICES_VE <- cbind(ICES_VE, CSquare2LonLat(ICES_VE$c_square, degrees=0.05))
spatBound <- list(xrange = range(ICES_VE$SI_LONG, na.rm=TRUE),
                  yrange = range(ICES_VE$SI_LATI, na.rm=TRUE))
spatCore  <- list(xrange = quantile(ICES_VE$SI_LONG, c(0.025, 0.975)),
                  yrange = quantile(ICES_VE$SI_LATI, c(0.025, 0.975)))
tempBound <- range(ICES_VE$year, na.rm=TRUE)
```
<!--------------------------
Checks for logbook data
------------------------ -->
```{r logbook-checks}
#Most recent submission -->
ICES_LE <- cbind(ICES_LE, ICESrectangle2LonLat(ICES_LE$ICES_rectangle, midpoint=TRUE))
spatBoundLog <- list(xrange = range(ICES_LE$SI_LONG, na.rm=TRUE),
                     yrange = range(ICES_LE$SI_LATI, na.rm=TRUE))
tempBoundLog <- range(ICES_LE$year, na.rm=TRUE)
```


<!--------------------------
Utility stuff
------------------------ -->
```{r ggplot settings}
theme_icesqc <- function(legend.position = "none")
  theme(axis.text.y   = element_text(colour="black"),
        axis.text.x   = element_text(colour="black"),
        axis.title.y  = element_text(size=14,face="bold"),
        axis.title.x  = element_text(size=14,face="bold"),
        legend.position = legend.position,
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA))
```

```{r utility functions}
make_polVMS <- function(coordGrd, resolution = 0.05) {
  polVMS <- data.frame(long = rep(c(-1, -1, 1,  1, -1), nrow(coordGrd)) * resolution/2,
                       lat  = rep(c(-1,  1, 1, -1, -1), nrow(coordGrd)) * resolution/2) 
  polVMS$long <- polVMS$long + rep(coordGrd$SI_LONG, each = 5)
  polVMS$lat <- polVMS$lat + rep(coordGrd$SI_LATI, each = 5)
  polVMS$order <- rep(1:5, nrow(coordGrd))
  polVMS$hole <- FALSE
  polVMS$piece <- 1
  polVMS$id <- rep(1:nrow(coordGrd), each = 5)
  polVMS$group <- paste0(polVMS$id, ".", polVMS$piece)
  
  polVMS
}
```

```{r default-spatial-ggplot}
spatialplot <- function(data, xyrange = spatCore) {
  ggplot(polLand, aes(long, lat)) +
    geom_polygon(aes(group=group), fill = "light grey") +
    coord_fixed(ratio = vmstools::lonLatRatio(min(xyrange$xrange), min(xyrange$yrange)),
                xlim = xyrange$xrange, 
                ylim = xyrange$yrange) +
    labs(x = "Longitude", y = "Latitude") +
    geom_polygon(data = data, 
                 aes(long, lat, group = group, fill = cols)) +
    geom_polygon(data = polLand, 
                 aes(long, lat, group = group), 
                 colour = "black", size = 0.25, fill = "transparent")
}
```

<!------------------------------------------------------------------------------
Tables & Figures VMS
---------------------------------------------------------------------------- -->

# Quality of the VMS data from `r country`
## Years & number of records for which data has been submitted:
```{r records-by-year, results='asis'}
kable(plyr::count(ICES_VE,'year'))

dat2plot <- data.frame(table(ICES_VE$year))
colnames(dat2plot) <- c("Year","Freq")

ggplot(dat2plot,aes(x=Year,y=Freq)) +
  geom_bar(stat="identity") + xlab("Year") + ylab("Number of records") +
  theme_icesqc()
```

\newpage

## Distribution of pings by month:
```{r pings-by-month}
kable(table(ICES_VE$month, ICES_VE$year))

qplot(factor(year),data=ICES_VE, geom="bar",fill=factor(month)) +
  xlab("Years") + ylab ("Count") + scale_fill_grey(guide=guide_legend(title="Month")) +
  theme_icesqc()
```

\newpage

## Spatial extent of data submitted by year:
```{r table-exent, results='asis'}
dat2tab <- as.matrix(aggregate(ICES_VE$SI_LONG,by=list(ICES_VE$year),FUN=range))
dat2tab <- cbind(dat2tab,as.matrix(aggregate(ICES_VE$SI_LATI,by=list(ICES_VE$year),FUN=range)[,-1]))
colnames(dat2tab) <- c("Year", "min lon", "max lon", "min lat", "max lat")
kable(dat2tab)
```

## Area for which data has been submitted:
```{r area-of-data,  fig.height = 9}
coordGrd <- unique(ICES_VE[, c("SI_LONG", "SI_LATI", "year")])

# create a fortied polygon of csquares
polVMS <- make_polVMS(coordGrd, resolution = 0.5)
polVMS$year <- rep(coordGrd$year, each = 5)
polVMS$cols <- "red"

spatialplot(polVMS, spatBound) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

```{r area-of-data2,  fig.height = 9}
coordGrd <- unique(ICES_VE[, c("SI_LONG", "SI_LATI", "year")])

# create a fortied polygon of csquares
polVMS <- make_polVMS(coordGrd)
polVMS$year <- rep(coordGrd$year, each = 5)
polVMS$cols <- "red"

spatialplot(polVMS, spatCore) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Frequency of vessel length categories by year:
```{r records-by-vessel-length, results='asis'}
kable(table(ICES_VE$vessel_length_category,ICES_VE$year))

qplot(factor(year), data = ICES_VE, geom = "bar", fill = factor(vessel_length_category)) +
  xlab("Years") + ylab ("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Length category")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Frequency of gear codes by year:
```{r records-by-gear-code, results='asis'}
kable(table(ICES_VE$gear_code,ICES_VE$year))

qplot(factor(year), data = ICES_VE, geom = "bar", fill = factor(gear_code)) +
  xlab("Years") + ylab ("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Gear code")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Spatial extend of 3 most dominant gears:
```{r spatial-extent-dominant-gears, fig.height = 9}
# get 3 main gears
gear_table <- with(ICES_VE, aggregate(fishing_hours, list(gear_code = gear_code), sum, na.rm = TRUE))
gear_table <- gear_table[order(gear_table$x, decreasing = TRUE),]
top3gears <- gear_table$gear_code[1:3]

# split by gear
idx <- which(ICES_VE$gear_code %in% top3gears)
coordGrd <- unique(ICES_VE[idx, c("SI_LONG", "SI_LATI", "year", "c_square", "gear_code")])

# create a fortied polygon of csquares : 0.39 secs
polVMS <- make_polVMS(coordGrd)
polVMS$year <- rep(coordGrd$year, each = 5)
polVMS$c_square <- rep(coordGrd$c_square,each=5)
polVMS$gear_code<- rep(coordGrd$gear_code,each=5)

# aggregate fishing days
dat <- with(ICES_VE[idx,],
            aggregate(fishing_hours/24,
                 by = list(year = year, c_square = c_square, gear_code = gear_code),
                 FUN = sum, na.rm = TRUE))
dat <- dplyr::rename(dat, fishing_days = x)

# legend calculations
steps <- ceiling(max(dat$fishing_days, na.rm = TRUE)/250)
breaks <- c(-1, 0, steps * c(1, 2.5, 5, 10, 25, 50, 100, 250))
legval <- paste(breaks[-length(breaks)], "<=", breaks[-1L])
legval[1] <- "0"

palette <- c("white", RColorBrewer::brewer.pal(length(breaks)-2, "YlOrRd"))  
dat$colgrp <- as.numeric(cut(dat$fishing_days, breaks = breaks))
dat$cols <- palette[dat$colgrp]

# join dat onto polVMS (much faster than left_join or merge)
# make joining keys
a_by <- apply(dat[c("year","c_square","gear_code")], 1, paste, collapse = ".")
b_by <- apply(polVMS[c("year","c_square","gear_code")], 1, paste, collapse = ".")
ib_by <- as.integer(factor(b_by, levels = a_by))
polVMS$cols <- dat$cols[ib_by]

# do plots
invisible(lapply(top3gears, function(gear) {
spatialplot(polVMS[polVMS$gear_code == gear,]) +
  scale_fill_manual(values = rev(palette), labels = rev(legval)) +
  guides(fill = guide_legend(title = paste("Days@Sea", gear))) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc(legend.position = "top")
}))
```



## Number of unique DCF Level 6 codes by year:
```{r table-met6, results='asis'}
kable(count(unique(ICES_VE[,c("LE_MET_level6","year")]),"year"))
```

Top 5 DCF Level 6 codes by year:
```{r top5-met6} 
top5Met6 <- names(sort(table(ICES_VE$LE_MET_level6), decreasing = TRUE)[1:5])
kable(table(ICES_VE$LE_MET_level6, ICES_VE$year)[top5Met6,])

qplot(factor(year), 
      data = ICES_VE[ICES_VE$LE_MET_level6 %in% top5Met6,],
      geom = "bar", fill = factor(LE_MET_level6)) +
  xlab("Years") + ylab ("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Level 6")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Average fishing speed:
```{r average-fishing-speed}
ggplot(ICES_VE, aes(x = avg_fishing_speed)) +
  geom_histogram() +
  xlab("Average fishing speed") + ylab ("Count") +
  facet_wrap( ~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Average fishing hours:
```{r fishing-hours}
ggplot(ICES_VE, aes(x = fishing_hours)) +
  geom_histogram() +
  xlab("Average fishing hours") + ylab("Count") +
  facet_wrap( ~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Average length:
```{r length}
ggplot(ICES_VE, aes(x = avg_oal))+
  geom_histogram() +
  xlab("Average vessel length") + ylab("Count") +
  facet_wrap( ~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Average kW:
```{r average-length}
ggplot(ICES_VE, aes(x = avg_kw))+
  geom_histogram() +
  xlab("Average kW") + ylab ("Count") +
  facet_wrap( ~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Average kW-hours:
```{r fishing-speed}
ggplot(ICES_VE, aes(x = kw_fishinghours))+
  geom_histogram() +
  xlab("Average kW-hours") + ylab("Count") +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Landings by gear by year:
```{r landings-by-gear-year}
dat2tab <- 
  with(ICES_VE, 
       tapply(totweight, list(gear_code = gear_code, year = year), sum, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "totweight")

ggplot(dat2plot, aes(x = year, y = totweight)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("kg landed") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Spatial distribution of effort by year:
```{r spatial-effort-year, results='asis', fig.height = 9}

coordGrd <- unique(ICES_VE[,c("SI_LONG","SI_LATI","year","c_square")])
# make fortified DF for csquares
polVMS <- make_polVMS(coordGrd)
polVMS$year     <- rep(coordGrd$year, each = 5)
polVMS$c_square <- rep(coordGrd$c_square, each = 5)

# aggregate fishing days
dat <- with(ICES_VE,
            aggregate(fishing_hours/24,
                 by = list(year = year, c_square = c_square),
                 FUN = sum, na.rm = TRUE))
dat <- dplyr::rename(dat, fishing_days = x)

# legend calculations
steps <- ceiling(max(dat$fishing_days, na.rm = TRUE)/250)
breaks <- c(-1, 0, steps * c(1, 2.5, 5, 10, 25, 50, 100, 250))
legval <- paste(breaks[-length(breaks)], "<=", breaks[-1L])
legval[1] <- "0"

palette <- c("white", RColorBrewer::brewer.pal(length(breaks)-2, "YlOrRd"))  
dat$colgrp <- as.numeric(cut(dat$fishing_days, breaks = breaks))
dat$cols <- palette[dat$colgrp]

# join dat onto polVMS (much faster than left_join or merge)
# make joining keys
a_by <- apply(dat[c("year","c_square")], 1, paste, collapse = ".")
b_by <- apply(polVMS[c("year","c_square")], 1, paste, collapse = ".")
ib_by <- as.integer(factor(b_by, levels = a_by))
polVMS$cols <- dat$cols[ib_by]

# do plots
spatialplot(polVMS) +
  scale_fill_manual(values = rev(palette), labels = rev(legval)) +
  guides(fill = guide_legend(title = "Days@Sea")) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc(legend.position = "top")
```

\newpage

## Spatial difference of effort `r tempBound[1]`:`r (tempBound[2]-1)` vs `r tempBound[2]`
```{r spatial-effort--difference1, results='asis', fig.height = 9}
base <- with(ICES_VE, 
             aggregate(fishing_hours, 
                       by = list(c_square = c_square, year = year),
                       FUN = sum, na.rm = TRUE))
base <- dplyr::rename(base, fishing_hours = x)

# calculate total fishing hours for recent year
recent <- base[base$year == tempBound[2],]

# calculate median of the total fishing hours per square for historical years
base <- with(base[base$year < tempBound[2],], 
             aggregate(fishing_hours,
                       by = list(c_square = c_square), 
                       FUN = median, na.rm = TRUE))
base <- dplyr::rename(base, fishing_hours_median = x)

# join
dat2plot <- dplyr::full_join(base,
                             recent[,c("c_square","fishing_hours")])

# set NAs to zero
dat2plot$fishing_hours_median[is.na(dat2plot$fishing_hours_median)] <- 0
dat2plot$fishing_hours[is.na(dat2plot$fishing_hours)] <- 0

# calculate ratio (with exceptions for zeros)
dat2plot$ratio <- with(dat2plot, pmax(fishing_hours_median, 1e-9) / pmax(fishing_hours, 1e-9))

# add back in lat and long
dat2plot <- cbind(dat2plot, 
                  vmstools::CSquare2LonLat(dat2plot$c_square, degrees = 0.05))

# make 'fortified' data frame
polVMS <- make_polVMS(dat2plot)
polVMS$c_square <- rep(dat2plot$c_square, each = 5)

# set up legend
# This is not what the legend says.... there is no +/-5% break!
breaks <- c(-Inf, 1/2, 1/1.5, 1/1.25, 1/1.05, 1, 1.05, 1.25, 1.5, 2, Inf)
legval <- c("historic >>","historic> +100%","historic> +50%","historic> +25%",
            "+/-5%",
            "recent> +5%","recent> +25%","recent> +50%","recent> +100%","recent >>")
palette <- RColorBrewer::brewer.pal(length(breaks)-1,"RdYlBu")  #colour for fishing intensity
dat2plot$colgrp <- as.numeric(cut(dat2plot$ratio, breaks = breaks))
dat2plot$cols <- palette[dat2plot$colgrp]

# join dat onto polVMS (much faster than left_join or merge)
# make joining keys
a_by <- dat2plot$c_square
b_by <- polVMS$c_square
ib_by <- as.integer(factor(b_by, levels = a_by))
polVMS$cols <- dat2plot$cols[ib_by]
# NOTE dark blue here just means that nothing was observered before, and now there is something... 
# - could be 1 hour of fishing
spatialplot(polVMS) +
  guides(fill = guide_legend(title = "Days@Sea")) +
  scale_fill_manual(values = rev(palette), labels = legval) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Spatial difference of effort `r tempBound[2]-1` vs `r tempBound[2]`

```{r spatial-effort--difference2, results='asis', fig.height = 9}
base <- with(ICES_VE, 
             aggregate(fishing_hours, 
                       by = list(c_square = c_square, year = year),
                       FUN = sum, na.rm = TRUE))
base <- dplyr::rename(base, fishing_hours = x)

# calculate total fishing hours for recent year
recent <- base[base$year == tempBound[2],]

# previous year
base <- base[base$year == tempBound[2]-1,]
base <- dplyr::rename(base, fishing_hours_median = fishing_hours)

# join
dat2plot <- dplyr::full_join(base,
                             recent[,c("c_square","fishing_hours")])

# set NAs to zero
dat2plot$fishing_hours_median[is.na(dat2plot$fishing_hours_median)] <- 0
dat2plot$fishing_hours[is.na(dat2plot$fishing_hours)] <- 0

# calculate ratio (with exceptions for zeros)
dat2plot$ratio <- with(dat2plot, pmax(fishing_hours_median, 1e-9) / pmax(fishing_hours, 1e-9))

# add back in lat and long
dat2plot <- cbind(dat2plot, 
                  vmstools::CSquare2LonLat(dat2plot$c_square, degrees = 0.05))

# make 'fortified' data frame
polVMS <- make_polVMS(dat2plot)
polVMS$c_square <- rep(dat2plot$c_square, each = 5)

# set up legend
# This is not what the legend says.... there is no +/-5% break!
breaks <- c(-Inf, 1/2, 1/1.5, 1/1.25, 1/1.05, 1, 1.05, 1.25, 1.5, 2, Inf)
legval <- c("historic >>","historic> +100%","historic> +50%","historic> +25%",
            "+/-5%",
            "recent> +5%","recent> +25%","recent> +50%","recent> +100%","recent >>")
palette <- RColorBrewer::brewer.pal(length(breaks)-1,"RdYlBu")  #colour for fishing intensity
dat2plot$colgrp <- as.numeric(cut(dat2plot$ratio, breaks = breaks))
dat2plot$cols <- palette[dat2plot$colgrp]

# join dat onto polVMS (much faster than left_join or merge)
# make joining keys
a_by <- dat2plot$c_square
b_by <- polVMS$c_square
ib_by <- as.integer(factor(b_by, levels = a_by))
polVMS$cols <- dat2plot$cols[ib_by]
# NOTE dark blue here just means that nothing was observered before, and now there is something... 
# - could be 1 hour of fishing
spatialplot(polVMS) +
  guides(fill = guide_legend(title = "Days@Sea")) +
  scale_fill_manual(values = rev(palette), labels = legval) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Mean landing per kW fishing hours by year:

```{r landings-by-gear}
dat2tab <- 
  with(ICES_VE, 
       tapply(totweight/kw_fishinghours, list(gear_code = gear_code, year = year), median, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "kgkWh")

ggplot(dat2plot, aes(x = year, y = kgkWh)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("kg/kWh") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Value by gear by year:
```{r value-by-gear-year, results='asis'}
dat2tab <- 
  with(ICES_VE, 
       tapply(totvalue, list(gear_code = gear_code, year = year), sum, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "EUR_landed")

ggplot(dat2plot, aes(x = year, y = EUR_landed)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("EUR landed") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Median value per KW fishing hours by year:

```{r mean-value-by-KWhours-year, results='asis'}
dat2tab <- 
  with(ICES_VE, 
       tapply(totvalue/kw_fishinghours, list(gear_code = gear_code, year = year), median, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "EURkWh")

ggplot(dat2plot, aes(x = year, y = EURkWh)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("EUR/kWh") +
  theme_icesqc(legend.position = "right")
```

\newpage

##  Average price:

```{r price, results='asis'}
dat2tab <- 
  with(ICES_VE, 
       tapply(totvalue/totweight, list(gear_code = gear_code, year = year), median, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "MeanPrice")

ggplot(dat2plot, aes(x = year, y = MeanPrice)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("Mean price") +
  theme_icesqc(legend.position = "right")
```

\newpage









<!--- 
                       --------------
                       LOGBOOK CHECKS
                       --------------
--->


# Quality of the logbook

## Years & number of records for which data has been submitted:
```{r logbook-records, results='asis'}
kable(count(ICES_LE,'year'))

dat2plot <- data.frame(table(ICES_LE$year))
colnames(dat2plot) <- c("Year","Freq")
ggplot(dat2plot,aes(x=Year,y=Freq)) +
  geom_bar(stat="identity") + xlab("Year") + ylab("Number of records") +
  theme_icesqc()
```

\newpage

## Distribution of logbook entries by month:
```{r logbook-entries-month, results='asis'}
kable(table(ICES_LE$month, ICES_LE$year))

qplot(factor(year),data=ICES_LE, geom="bar",fill=factor(month)) +
  xlab("Years") + ylab ("Count") + scale_fill_grey(guide=guide_legend(title="Month")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Area extent of data submitted by year:
```{r , results='asis', fig.height = 9}
dat2tab <- as.matrix(aggregate(ICES_LE$SI_LONG, by = list(year = ICES_LE$year), FUN=range, na.rm=TRUE))
dat2tab <- cbind(dat2tab,as.matrix(aggregate(ICES_LE$SI_LATI,by=list(ICES_LE$year),FUN=range)[,-1]))
colnames(dat2tab) <- c("Year","min lon","max lon","min lat","max lat")
kable(dat2tab)
```

## Area for which data has been submitted:
```{r , results='asis',fig.height = 9}
coordGrd        <- unique(ICES_LE[,c("SI_LONG","SI_LATI","year")])

polRect <- make_polVMS(coordGrd, resolution = 1)
polRect$year    <- rep(coordGrd$year, each = 5)
polRect$cols    <- "red"

spatialplot(polRect, spatBoundLog) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```


```{r , results='asis',fig.height = 9}
spatialplot(polRect, spatCore) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Frequency of vessel length categories by year:
```{r , results='asis'}
kable(table(ICES_LE$vessel_length_category,ICES_LE$year))

qplot(factor(year), data = ICES_LE, geom = "bar", fill = factor(vessel_length_category)) +
  xlab("Years") + ylab ("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Length category")) +
  theme_icesqc()
```

\newpage

## Frequency of gear codes by year:
```{r , results='asis'}
kable(table(ICES_LE$gear_code,ICES_LE$year))

qplot(factor(year), data = ICES_LE, geom = "bar", fill = factor(gear_code)) +
  xlab("Years") + ylab("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Gear code")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Number of unique DCF Level 6 codes by year:
```{r , results='asis'}
kable(count(unique(ICES_LE[,c("LE_MET_level6","year")]),"year"))
```

## Top 5 DCF Level 6 codes by year:
```{r , results='asis'}
top5Met6 <- names(rev(sort(table(ICES_LE$LE_MET_level6)))[1:5])
kable(table(ICES_LE$LE_MET_level6,ICES_LE$year)[top5Met6,])

qplot(factor(year), data = subset(ICES_LE,LE_MET_level6 %in% top5Met6),
      geom = "bar", fill = factor(LE_MET_level6)) +
  xlab("Years") + ylab("Count") + 
  scale_fill_grey(guide = guide_legend(title = "Level 6")) +
  theme_icesqc(legend.position = "right")
```

\newpage

## Frequency of VMS enabled category
```{r , results='asis'}
kable(table(ICES_LE$vms_enabled,ICES_LE$vessel_length_category))
```

## Average fishing days:
```{r , results='asis'}
ggplot(ICES_LE, aes(x = fishing_days))+
  geom_histogram() +
  xlab("Average fishing days") + ylab("Count") +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Average KW-hours:
```{r , results='asis'}
ggplot(ICES_LE, aes(x = kw_fishing_days)) +
  geom_histogram() +
  xlab("Average KW-days") + ylab("Count") +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc()
```

\newpage

## Landings by gear by year:
```{r , results='asis'}
dat2tab <- 
  with(ICES_LE, 
       tapply(totweight, list(gear_code = gear_code, year = year), sum, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "totweight")

ggplot(dat2plot, aes(x = year, y = totweight)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd = 1.5) +
  xlab("Year") + ylab ("kg landed") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Spatial distribution of effort by year:
```{r , results='asis', fig.height = 9}
coordGrd        <- unique(ICES_LE[,c("SI_LONG","SI_LATI","year","ICES_rectangle")])

polRect  <- make_polVMS(coordGrd, resolution = 1)
polRect$year    <- rep(coordGrd$year,each=5)
polRect$ICES_rectangle <- rep(coordGrd$ICES_rectangle, each = 5)

#TIDY ---
dat             <- aggregate(ICES_LE$fishing_days,by=list(ICES_LE$year,ICES_LE$ICES_rectangle),FUN=sum,na.rm=T)

steps               <- ceiling(max(dat$x,na.rm=T)/250)
cutbreaksval        <- c(-1,0,steps*c(1,2.5,5,10,25,50,100,250))
legval              <- outer(ac(cutbreaksval),ac(cutbreaksval),function(x,y){return(paste(x,"<=",y))})
legval              <- c("0",diag(legval[-1,-c(1,2)]))
palette <- c("white", brewer.pal(length(cutbreaksval)-2,"YlOrRd"))
cols <- palette[cut(dat$x,breaks=cutbreaksval)]
cols                <- cbind(cols,id=1:length(cols),ICES_rectangle=dat$Group.2,year=dat$Group.1)
polRect              <- merge(polRect,cols,by=c("ICES_rectangle","year"))
# ----

spatialplot(polRect) +
  guides(fill = guide_legend(title = "Days@Sea")) +
  scale_fill_manual(values = rev(palette), labels = rev(legval)) +
  facet_wrap(~ year, ncol = 2) +
  theme_icesqc(legend.position = "top")
```

\newpage

## Spatial difference of effort `r tempBound[1]`:`r (tempBound[2]-1)` vs `r tempBound[2]`
```{r,  fig.height = 9}
base <- with(ICES_LE, 
             aggregate(fishing_days, 
                       by = list(ICES_rectangle = ICES_rectangle, year = year),
                       FUN = sum, na.rm = TRUE))
base <- dplyr::rename(base, fishing_days = x)

# calculate total fishing hours for recent year
recent <- base[base$year == tempBound[2],]

# calculate median of the total fishing hours per square for historical years
base <- with(base[base$year < tempBound[2],], 
             aggregate(fishing_days,
                       by = list(ICES_rectangle = ICES_rectangle), 
                       FUN = median, na.rm = TRUE))
base <- dplyr::rename(base, fishing_days_median = x)

# join
dat2plot <- dplyr::full_join(base,
                             recent[,c("ICES_rectangle","fishing_days")])

# set NAs to zero
dat2plot$fishing_days_median[is.na(dat2plot$fishing_days_median)] <- 0
dat2plot$fishing_days[is.na(dat2plot$fishing_days)] <- 0

# calculate ratio (with exceptions for zeros)
dat2plot$ratio <- with(dat2plot, pmax(fishing_days_median, 1e-9) / pmax(fishing_days, 1e-9))

# add back in lat and long
dat2plot <- cbind(dat2plot,
                  vmstools::ICESrectangle2LonLat(dat2plot$ICES_rectangle, midpoint = TRUE))

# make 'fortified' data frame
polRect <- make_polVMS(dat2plot, resolution = 1)
polRect$ICES_rectangle <- rep(dat2plot$ICES_rectangle, each = 5)

## tidy ---
breaks <- rev(c(1e-10,0.5,2/3,0.8,0.952381,1,1.05,1.25,1.5,2,1e10))
legval <- c("historic >>","historic> +100%","historic> +50%","historic> +25%",
            "+/-5%",
            "recent> +5%","recent> +25%","recent> +50%","recent> +100%","recent >>")
palette <- brewer.pal(length(cutbreaksval)-1,"RdYlBu")
colgrp <- as.numeric(cut(dat2plot$ratio, breaks = breaks))
cols <- cbind(cols = palette[colgrp], ICES_rectangle = dat2plot$ICES_rectangle)
polRect <- merge(polRect, cols, by=c("ICES_rectangle"))
# ---

spatialplot(polRect) +
    guides(fill=guide_legend(title="Days@Sea")) +
    scale_fill_manual(values = rev(palette), labels = legval) +
    theme_icesqc(legend.position = "right")
```

\newpage

## Spatial difference of effort `r tempBound[2]-1` vs `r tempBound[2]`

```{r,  fig.height = 9}
base <- with(ICES_LE, 
             aggregate(fishing_days, 
                       by = list(ICES_rectangle = ICES_rectangle, year = year),
                       FUN = sum, na.rm = TRUE))
base <- dplyr::rename(base, fishing_days = x)

# calculate total fishing hours for recent year
recent <- base[base$year == tempBound[2],]

# previous year
base <- base[base$year == tempBound[2]-1,]
base <- dplyr::rename(base, fishing_days_median = fishing_days)

# join
dat2plot <- dplyr::full_join(base,
                             recent[,c("ICES_rectangle","fishing_days")])

# set NAs to zero
dat2plot$fishing_days_median[is.na(dat2plot$fishing_days_median)] <- 0
dat2plot$fishing_days[is.na(dat2plot$fishing_days)] <- 0

# calculate ratio (with exceptions for zeros)
dat2plot$ratio <- with(dat2plot, pmax(fishing_days_median, 1e-9) / pmax(fishing_days, 1e-9))

# add back in lat and long
dat2plot <- cbind(dat2plot,
                  vmstools::ICESrectangle2LonLat(dat2plot$ICES_rectangle, midpoint = TRUE))

# make 'fortified' data frame
polRect <- make_polVMS(dat2plot, resolution = 1)
polRect$ICES_rectangle <- rep(dat2plot$ICES_rectangle, each = 5)

## tidy ---
breaks <- rev(c(1e-10,0.5,2/3,0.8,0.952381,1,1.05,1.25,1.5,2,1e10))
legval <- c("historic >>","historic> +100%","historic> +50%","historic> +25%",
            "+/-5%",
            "recent> +5%","recent> +25%","recent> +50%","recent> +100%","recent >>")
palette <- brewer.pal(length(cutbreaksval)-1,"RdYlBu")
colgrp <- as.numeric(cut(dat2plot$ratio, breaks = breaks))
cols <- cbind(cols = palette[colgrp], ICES_rectangle = dat2plot$ICES_rectangle)
polRect <- merge(polRect, cols, by=c("ICES_rectangle"))
# ---

spatialplot(polRect) +
    guides(fill=guide_legend(title="Days@Sea")) +
    scale_fill_manual(values = rev(palette), labels = legval) +
    theme_icesqc(legend.position = "right")
```

\newpage


## Relationship fishing days and total weight
```{r , fig.height = 9}
ggplot(ICES_LE[ICES_LE$year == tempBoundLog[2],], 
       aes(x = fishing_days, y = totweight)) +
  geom_point() +
  facet_wrap(~ gear_code, ncol = 3, scale = "free") +
  xlab("Fishing days") + ylab ("Total weight") +
  theme_icesqc()
```

\newpage

## Mean landing per KW fishing day by year:

```{r , results='asis'}
dat2tab <- 
  with(ICES_LE, 
       tapply(totweight/kw_fishing_days, list(gear_code = gear_code, year = year), median, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "kgkWh")

ggplot(dat2plot, aes(x = year, y = kgkWh)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("kg/kWh") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Value by gear by year:
```{r , results='asis'}
dat2tab <- 
  with(ICES_LE, 
       tapply(totvalue, list(gear_code = gear_code, year = year), sum, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "EUR_landed")

ggplot(dat2plot, aes(x = year, y = EUR_landed)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("EUR landed") +
  theme_icesqc(legend.position = "right")
```

\newpage

## Mean value per KW fishing day by year:

```{r , results='asis'}
dat2tab <- 
  with(ICES_LE, 
       tapply(totvalue/kw_fishing_days, list(gear_code = gear_code, year = year), median, na.rm = TRUE))
kable(dat2tab)

dat2plot <- as.data.frame.table(dat2tab, responseName = "EURkWh")

ggplot(dat2plot, aes(x = year, y = EURkWh)) +
  geom_line(aes(group = gear_code, colour = gear_code), lwd=1.5) +
  xlab("Year") + ylab ("EUR/kWh") +
  theme_icesqc(legend.position = "right")
```

\newpage

##  Average price:

```{r , results='asis'}
dat2tab <- 
  with(ICES_LE, 
       tapply(totvalue/totweight, list(year = year), median, na.rm = TRUE))
kable(as.data.frame(t(dat2tab)))

dat2plot <- as.data.frame.table(dat2tab, responseName = "MeanPrice")

ggplot(dat2plot, aes(x = year, y = MeanPrice, group = 1)) +
  geom_line(lwd = 1.5) +
  xlab("Year") + ylab ("Mean price") +
  theme_icesqc()
```

\newpage

## Comparison of Metier level 6 reporting between logbook and VMS
```{r met6-comparison, results = 'asis'}

ledat <- with(ICES_LE, table(LE_MET_level6, year))
vedat <- with(ICES_VE, table(LE_MET_level6, year))

dat2tab <- 
  rbind(
    cbind(as.data.frame.table(ledat), data = "LE (records)"),
    cbind(as.data.frame.table(vedat), data = "VE (pings)"))

tab <- with(dat2tab, tapply(Freq, list(LE_MET_level6, data, year), sum))
tab[tab == 0] <- NA

for (i in dimnames(tab)[[3]]) {
  x <- tab[,,i]
  x <- x[apply(!is.na(x), 1, any),]
  cat(kable(cbind(x, year = i), table.attr='class="flat-table"'), sep = "\n")
  cat("\n")
}
```